Cluster Analysis: Identifying Parkinson's Disease Subtypes

### 2015-06-26 writeup/parkinsons.tex
rerun preprocessing graphs, pca
rerun determining k graphs, png -> pdf
update kmeans statistics after rerun

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.76&  3.32&   0-24 \\
nms\_d2&   8.71&  8.76&   0-48 \\
nms\_d3&   8.70& 11.83&   0-60 \\
nms\_d4&   1.65&  3.94&   0-33 \\
nms\_d5&   5.22&  7.44&   0-36 \\
nms\_d6&   5.67&  6.92&   0-36 \\
nms\_d7&   8.02&  9.09&   0-36 \\
nms\_d8&   3.57&  5.97&   0-24 \\
nms\_d9&   6.99&  7.74&   0-48 \\
tremor&   2.59&  2.63&   0-12 \\
bradykin& 2.49&  1.39&   0-6 \\
rigidity& 2.34&  1.36&   0-6 \\
axial&    3.28&  2.75&   0-12 \\
pigd&     3.36&  2.77&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

May not be useful? If we're trying to identify \emph{clinically} relevant
features, merging them may not be a good idea.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{pca-eigenvalues.png}
  \caption{Scree test: eigenvalues by factor}
  \label{fig:pca-eigenvalues}
\end{figure}

Figure~\ref{fig:pca-eigenvalues} shows scree test elbow occurs around 2 or 3. Also, eigenvalues 1 and $2 > 1$, while 3
is around .9

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-wss-error.png}
  \caption{Scree test: WSS error by cluster size}
  \label{fig:kmeans-wss-error}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.png}
  \caption{Gap statistic by cluster size}
  \label{fig:gap-statistic}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{asw.png}
  \caption{Average silhouette width by cluster size}
  \label{fig:asw}
\end{figure}

@@

\subsection{Cluster statistics}
% CLUSTERS: 2
% ================================
% Sizes: 201 580
% WithinSS: 4248.585 4132.434
% Sum WithinSS: 8381.019
% CLUSTERS: 3
% ================================
% Sizes: 420 231 130
% WithinSS: 2618.368 1973.82 3076.542
% Sum WithinSS: 7668.73
% CLUSTERS: 4
% ================================
% Sizes: 61 372 145 203
% WithinSS: 1481.25 1845.389 2147.988 1609.555
% Sum WithinSS: 7084.183
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    $k$ & $n$ & Within SS & sum(Within SS) \\
    \hline
    2 & 201/580 & 4248.585/4132.434 & 8381.019 \\
    3 & 420/231/130 & 2618.368/1973.82/3076.542 &  7668.73 \\
    4 & 61/372/145/203 & 1481.25/1845.389/2147.988/1609.555 & 7084.183
  \end{tabular}
  \caption{Cluster statistics}
  \label{tab:cluster-statistics}
\end{table}

@@

\subsection{Decision trees based on clusters}
% seed = 911
% CLUSTERS: 2
% ================================
% Complexity Parameter: 0.03482587
% 10-fold CV error: 0.134443
% Root node error: 0.2573624
% CLUSTERS: 3
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.1946223
% Root node error: 0.4622279
% CLUSTERS: 4
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.2483995
% Root node error: 0.5236876
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l|l|l}
    $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    validation} & Root Feature &
    Root Error & Figure \\
    \hline
    2 & 0.0348 & 0.134 & axial $\geq$ 0.44 & 0.257 & Figure~\ref{fig:kmeans-dtree-2} \\
    3 & 0.0100 & 0.194 & bradykin $<$ 0.0041 & 0.462 & Figure~\ref{fig:kmeans-dtree-3} \\
    4 & 0.0100 & 0.248 & bradykin $<$ 0.0041 & 0.523 & Figure~\ref{fig:kmeans-dtree-4} \\
  \end{tabular}
  \caption{$k$-kmeans decision trees statistics}
  \label{tab:k-means-dtrees}
\end{table}

>>>>

\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
  Name  &       $\mu$ & $\sigma$ & min-max \\
         \hline
nms\_d1&   1.73&  3.35&   0-24 \\
nms\_d2&   8.75&  8.70&   0-48 \\
nms\_d3&   8.68& 11.55&   0-60 \\
nms\_d4&   1.64&  3.86&   0-33 \\
nms\_d5&   5.42&  7.43&   0-36 \\
nms\_d6&   5.53&  6.79&   0-36 \\
nms\_d7&   8.08&  8.94&   0-36 \\
nms\_d8&   3.52&  5.97&   0-24 \\
nms\_d9&   7.13&  7.79&   0-48 \\
tremor&   2.59&  2.58&   0-12 \\
bradykin& 2.40&  1.41&   0-6 \\
rigidity& 2.24&  1.36&   0-6 \\
axial&    3.25&  2.68&   0-12 \\
pigd&     3.31&  2.71&   0-12 \\
  \end{tabular}
  \caption{Descriptive Statistics}
  \label{tab:descriptive-statistics}
\end{table}

@@

May not be useful? If we're trying to identify \emph{clinically} relevant
features, merging them may not be a good idea. Regardless, Figure~\ref{fig:pca}
shows results of preliminary PCA.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{pca.pdf}
  \caption{PCA Analysis}
  \label{fig:pca}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{pca-eigenvalues.pdf}
  \caption{Scree test: eigenvalues by factor}
  \label{fig:pca-eigenvalues}
\end{figure}

Figure~\ref{fig:pca-eigenvalues} shows scree test elbow occurs around 2 or 2 or
.4 Also, eigenvalues $1-5 > 1$.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{kmeans-wss-error.pdf}
  \caption{Scree test: WSS error by cluster size}
  \label{fig:kmeans-wss-error}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{gap-statistic.pdf}
  \caption{Gap statistic by cluster size}
  \label{fig:gap-statistic}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{asw.pdf}
  \caption{Average silhouette width by cluster size}
  \label{fig:asw}
\end{figure}

@@

\subsection{Cluster statistics}
% CLUSTERS: 2
% ================================
% Sizes: 229 672
% WithinSS: 6117.986 7695.411
% Sum WithinSS: 13813.4
% CLUSTERS: 3
% ================================
% Sizes: 333 134 434
% WithinSS: 4669.037 4009.387 4153.892
% Sum WithinSS: 12832.32
% CLUSTERS: 4
% ================================
% Sizes: 79 394 275 153
% WithinSS: 2366.585 3356.709 3454.142 2879.508
% Sum WithinSS: 12056.94
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l}
    $k$ & $n$ & Within SS & sum(Within SS) \\
    \hline
    2 & 229/672 & 6118/7695 & 13813 \\
    3 & 333/134/434 & 4669/40009/4154 & 12832 \\
    4 & 79/394/275/153 & 2367/3357/3454/2880  & 12057 \\
  \end{tabular}
  \caption{Cluster statistics}
  \label{tab:cluster-statistics}
\end{table}

@@

\subsection{Decision trees based on clusters}
% seed = 911
% CLUSTERS: 2
% ================================
% Complexity Parameter: 0.02183406
% 10-fold CV error: 0.1132075
% Root node error: 0.254162
% CLUSTERS: 3
% ================================
% Complexity Parameter: 0.01070664
% 10-fold CV error: 0.190899
% Root node error: 0.518313
% CLUSTERS: 4
% ================================
% Complexity Parameter: 0.01
% 10-fold CV error: 0.2552719
% Root node error: 0.5627081
\begin{table}[h]
  \centering
  \begin{tabular}{l|l|l|l|l|l}
    $k$ & CP\tablefootnote{Complexity Parameter} & CV Xerror\tablefootnote{10-fold cross
    validation} & Root Feature &
    Root Error & Figure \\
    \hline
    2 & 0.0218 & 0.113 & axial $\geq$ 4.5 & 0.254 & Figure~\ref{fig:kmeans-dtree-2} \\
    3 & 0.0107 & 0.191 & pigd $\geq$ 2.5 & 0.518 & Figure~\ref{fig:kmeans-dtree-3} \\
    4 & 0.0100 & 0.255 & pigd $<$ 2.5 & 0.563 & Figure~\ref{fig:kmeans-dtree-4} \\
  \end{tabular}
  \caption{$k$-kmeans decision trees statistics}
  \label{tab:k-means-dtrees}
\end{table}
